\documentclass{easychair}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{graphicx}

\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\title{Constructing QITs from Quotients}

\author{
    Thorsten Altenkirch\inst{1}
\and
    Christina O'Donnell\inst{1}\thanks{Supported by the Engineering and Physical Sciences Research Council (EPSRC) Doctoral Landscape Award [Grant Ref: EP/Z534948/1].}
}

\institute{
    University of Nottingham, UK\\
    \email{\{psztxa,psxco3\}@nottingham.ac.uk}
}

\authorrunning{T. Altenkirch and C. D. O'Donnell}
\titlerunning{Constructing QITs from Quotients}

\begin{document}

\maketitle

\section{Introduction}

Quotient inductive types (QITs) were introduced as a set-level
fragment of higher inductive types suitable for internalising syntax
in type theory~\cite{altenkirch2016-qit}. In a UIP or set-truncated
setting, the higher structure collapses, but the idea of specifying
data together with identifications remains, leading to quotient
inductive types with initial-algebra semantics and elimination by
initiality rather than by quotienting raw syntax
afterwards~\cite{altenkirch2018-qiit,kaposi2019-constructing,kovacs2020-large}. This
talk focuses on a basic obstruction in that picture: although a QIT
may at first seem to be just an ordinary inductive type quotiented by
equations, extensional quotienting forgets information that
constructors may still need. This tension becomes especially visible
in the infinitary case, where known general constructions appeal to
weak choice principles. Using a simple running example, I explain why
this happens and why weak quotients and setoid-style presentations
suggest a different constructive moral.

\section{Constructing QITs}

A quotient inductive type is not
merely a quotient of raw terms, but an initial algebra for a signature
with equations. If $F_\Sigma(X)$ is the polynomial functor associated
to a signature $\Sigma = (S,P)$, then a candidate quotient inductive
type is given by an algebra
\[
\mathsf{sup} : F_\Sigma(Q) \to Q
\]
together with the expected universal property. The point of this
presentation is that it packages the constructors and the eliminator at
the same time: what matters is not only which terms are identified, but
also how the resulting object supports recursion and induction. This is
the perspective developed in the semantic treatment of QIITs by
Altenkirch, Capriotti, Dijkstra, Kraus and Nordvall
Forsberg~\cite{altenkirch2018-qiit}, and later extended to finitary and
large/infinitary signature languages by Kaposi, Kov\'acs and
collaborators~\cite{kaposi2019-constructing,kovacs2020-large}.

\section{The lifting problem}

The central example of the talk is the type of \emph{mobiles}: countably
branching trees in which permuting the children does not change the
resulting element. Writing $M$ for the desired quotient type, one may
present it schematically by constructors
\[
\ell : M,
\qquad
n : (\mathbb{N}\to M)\to M,
\qquad
\mathsf{perm}_\pi(f) : n(f\circ \pi)=n(f)
\]
for each permutation $\pi : \mathrm{Perm}(\mathbb{N})$.

The naive construction starts from the raw $W$-type $W$ of ordered
countably branching trees and forms a quotient $W/{\sim}$ by the
permutation equations. At first glance this looks exactly right. But the
problem appears as soon as we ask for the constructor on the quotient.
The leaf constructor lifts immediately by sending $\ell_W$ to its
equivalence class. For the node constructor, however, we would need to
define
\[
n_{\sim} : (W/\sim)^{\mathbb{N}} \to W/\sim.
\]
The obvious route is: given a family of quotient elements, choose a raw
tree representing each one, apply the raw constructor in $W$, and then
quotient again. But this requires choosing representatives.

This is the basic lifting problem:

\begin{quote}
a quotient gives equivalence classes, but a constructor wants
representatives.
\end{quote}

That is the conceptual centre of the talk. It explains why quotient
classes alone are not enough: a QIT must still support elimination, and
that means the constructors must be well defined on quotient elements,
not just on chosen representatives.

\section{Why the problem is choice-shaped}

This obstruction already has the shape of a choice principle. If the
quotient map $[-]_\sim : W \twoheadrightarrow W/\sim$ admitted a section, then
we could pick a representative tree for each quotient element and define
$n_M$ by lifting along that section. In $\mathrm{Set}$, the
principle that every surjection splits is equivalent to the axiom of
choice. Thus the difficulty is not accidental: choice appears precisely
when extensional quotienting has forgotten data that the constructor
still needs.

In the infinitary case this phenomenon is reflected in existing general
constructions. Fiore, Pitts and Steenkamp introduce a broad class of
indexed quotient-inductive types, called QWI types, and show that they
can be constructed from inductive types and quotient types under WISC
(or, equivalently for the relevant argument, IWISC in the form used in
their paper)~\cite{fiore2022-qwit}. Their construction proceeds
stagewise, building size-bounded approximations and then taking a
colimit. The hard step is showing that forming constructor inputs is
compatible with this stagewise colimit; this is exactly where the weak
choice principle enters.

This fits a wider semantic picture. Higher inductive types are designed
to build identifications into the type former itself rather than by
external quotienting, and Lumsdaine and Shulman give general semantics
for such constructions in suitable model categories~\cite{lumsdaine2020-hit}.
From the present viewpoint, the contrast is not that one setting has
problems and the other does not, but that different equality formalisms
forget different amounts of information.

\section{Choosing not to forget}

The second half of the talk takes this as a design hint. If choice
appears because extensional quotients forget too much, then perhaps a
more constructive notion of quotient should forget less. This motivates
weak quotients and setoid-style presentations, where points are retained
and equality is carried as explicit structure rather than collapsed
immediately to bare equivalence classes. On this view, one does not need
to recover representatives later, because the witness data was never
thrown away in the first place.

This does \emph{not} mean that weak quotients solve everything. They
solve one problem by introducing another. In a setoid or displayed-setoid
semantics, a dependent family over a context must not only assign fibres
to points, but also provide transport maps along equality witnesses,
together with the required coherence conditions. So the difficulty moves
from choosing representatives to managing coherent transport. Kaposi and
Xie's work on quotient inductive-inductive types in the setoid model is
a useful reference point here: it shows how much can be recovered in a
strict setoid semantics, but also makes clear that plain setoids do not
automatically provide all the coherence one would like for dependent
elimination~\cite{kaposi2021-setoid}.

\section{Conclusion}

The aim of the talk is not to present a completed new construction,
but to sharpen the conceptual picture. Using mobiles as a simple
infinitary example, I argue that quotient inductive types are
difficult because extensional quotients erase witness data and then
require something choice-like when that data must later be
recovered. The stagewise construction of Fiore, Pitts and Steenkamp
shows how far this can be repaired using weak choice, while weak
quotients suggest a different constructive moral: do not forget the
witness data too early. The price, however, is that dependent type
theory must then carry coherent transport explicitly. In this sense,
the real issue is not whether we quotient, but what information our
notion of equality is allowed to forget.

\section{Rev1}
\section{Introduction and Motivation}\label{introduction-and-motivation}

Quotient Inductive Types (QITs) are a central mechanism for specifying
datatypes equipped with identifications in set-truncated Homotopy Type
Theory \cite{hott2013-book}. They support a wide range of fundamental
constructions, including the HoTT reals \cite{hott2013-book}, partiality
monads \cite{altenkirch2017-partiality-monad}, and ordinal-like
structures \footnote{TODO: Citaiton}.

Despite their expressive power, QITs present significant foundational
difficulties. Shulman and Lumsdaine established a no-go theorem showing
that certain QITs, including Brower's definition of the countable ordinals,
cannot be constructed using quotients alone
\cite{lumsdaine2020-semantics-higher-inductive}. This suggests that some
form of infinitary principle or choice is unavoidable in a constructive
metatheory.

Fiore, Pitts, and Steenkamp subsequently showed that a broad class of
QITs, which they call Quotient-W Types (QW types), admit an Initial
Algebra semantics assuming the Weak Initial Set of Covers (WISC)
principle \cite{fiore2022-quotients-inductive-types}. Their result
recovers many important examples, including the extensional countable
ordinals. However, WISC remains a nontrivial choice principle
\cite{vandenberg2012-wisc-axiom}, and its necessity is poorly
understood. In particular, it is unclear which features of a QIT
genuinely require choice, and which arise from limitations of existing
proof techniques.

In this work we isolate two structural properties of quotient systems
that explains this distinction: \emph{locality} and the tighter property
of \emph{depth-delta}. Intuitively, locality measures how deeply one
must inspect the inductive structure of terms in order to witness that
two elements are equal while depth-delta measures the maximum depth
change an equation. In pathological examples, such as tree ordinals, the
principle of extensionality quantifies over \emph{all} ordinals less
than a pair of ordinals, unbounded descent into subtrees, forcing the
use of global choice principles. By contrast, in examples such as
infinite mobiles---infinitely branching trees quotiented by
permutation---equality preserves depth and never requires inspecting
substructure beyond a fixed finite level. In fact because permutation
doesn't change the subtree relation, we can say that it is in a class of
depth-preserving QW types, along with the classical definition of
multi-sets as a QW type over a list.

Our key observation is that this boundedness of inspection depth is
exactly what is needed to recover the Fiore--Pitts--Steenkamp
construction constructively. We sketch a proof that for QW signatures
whose equations are bounded by a fixed ordinal \(\alpha\), cocontinuity
of the associated polynomial functor can be proved without any appeal to
any choice principle. As a consequence, we obtain a fully constructive
construction of infinite mobiles as an initial QIT, contrary to previous
expectations, which we have formalized in Agda
\cite{vezzosi2021-cubical-agda-dependently}.

This leads to our central research question:

\begin{quote}
	Which quotient-W types can be constructed fully constructively, without
	assuming any form of choice?
\end{quote}

We propose bounded locality as a criterion separating constructible QITs
from genuinely non-constructive ones, and we develop a framework for
expressing and exploiting this boundedness in the construction of
initial QW-algebras.

\section{Background}\label{background}

We briefly recall the framework of quotient-W types introduced by Fiore,
Pitts, and Steenkamp \cite{fiore2022-quotients-inductive-types}. A QW
signature consists of:

\begin{itemize}
	\item
	      A container \(\Sigma = S \triangleleft P\)
	      \cite{abbott2005-containers}, specifying the shape and positions of a
	      Polynomial Functor whose initial algebra is a W-Type
	      \cite{martinlof1984-intuitionistic}.
	\item
	      A system of equations \((E, V, l, r)\), where each equation \(e : E\)
	      has a set of variables \(V(e)\) and left- and right-hand sides given
	      by terms in the free \(\Sigma\)-algebra over those variables.
\end{itemize}

From such a signature, Fiore et al.~construct a diagram of
approximations indexed by a size structure
\cite{fiore2022-quotients-inductive-types}. Intuitively, this diagram
represents successive stages of the quotient construction, where each
stage gives the proofs available within a bounded depth. The desired QW
type is obtained as the Colimit of this diagram.

To show that this colimit yields an initial algebra, two key properties
are required:

\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.}
	\tightlist
	\item
	      The existence of suitable covers for the underlying W-type and the
	      equation contexts.
	\item
	      \emph{Cocontinuity} of the associated polynomial functor,
	      i.e.~preservation of the colimit.
\end{enumerate}

Both properties are established in Fiore et al.~using the indexed WISC
principle \cite{fiore2022-quotients-inductive-types}. In particular,
WISC is crucial in proving cocontinuity, which amounts to constructing a
isomorphism for the functor \(F\) determined by the container
\(\Sigma\).

Operationally, cocontinuity expresses the fact that applying
constructors commutes with taking quotients: forming a node after
quotienting is equivalent to quotienting after forming a node. The proof
in Fiore et al.~relies on WISC to select a cover to size-bound the
family of subtrees, enabling the construction of inverse maps witnessing
this isomorphism.

Our work revisits this step. We observe that the need for WISC arises
precisely when the equations allow arbitrarily deep inspection of terms,
as in the extensional ordinal example. When all equations are
depth-bounded, the required coherence data can instead be constructed by
well-founded recursion on a fixed ordinal bound \(\alpha\), avoiding any
appeal to choice.

\section{Locality Principle}\label{locality-principle}

We formalise the notion of locality by assigning a \emph{rank} function

\[ \mathrm{rank} : T_\Sigma X \to \mathcal{O} \]

measuring the height of the underlying tree. An equation system is said
to be \(\alpha\)-local if, for every equation \(e : E\), both sides
\(l(e)\) and \(r(e)\) have rank \emph{at most} a fixed ordinal
\(\alpha\). Intuitively, this means that every generating equation only
inspects structure up to depth \(\le \alpha\).

\[ \mathrm{rank} : T_\Sigma X \to \mathcal{O} \]

\section{Main Result}\label{main-result}

Our main theorem is that \(\alpha\)-local QW signatures admit fully
constructive initial algebra semantics.

\emph{Theorem} Let \(\Sigma\) be a QW signature whose equations are
\(\alpha\)-local. Then the associated quotient-W type exists as an
initial algebra, and its construction does not require WISC or any other
choice principle.

The proof follows the outline of Fiore--Pitts--Steenkamp
\cite{fiore2022-quotients-inductive-types}, but replaces the use of WISC
in the inverse map by offsetting by a bound ordinal \(\alpha\), giving a
constructable function between term bound and maximum proof depth,
giving cocontinuity. This construction is fully constructive and does
not require any choice principles.

\section{Case Study: Infinite
  Mobiles}\label{case-study-infinite-mobiles}

Our formalisation in Agda proves that in the case of mobiles, the
defining equations are provably \emph{depth-preserving}, since all
variables on the left appear at the same depth on the right. This is
enough to show that depth is preserved in the quotient type, and that if
two terms are equal then they can be proven in the stage bounded by
either ordinal.

As a result, the main theorem yields a fully constructive construction
of infinite mobiles as an initial QIT. This provides a concrete example
of a genuinely infinitary quotient type that does \emph{not} require
choice, contrary to previous assumptions.

\section{Boundary Phenomena}\label{boundary-phenomena}

Our analysis explains why certain examples genuinely require choice. In
the case of extensional ordinals, equality is defined by mutual
containment of downward closures, which requires unbounded descent into
subtrees \cite{fiore2022-quotients-inductive-types}. No ordinal bound
\(\alpha\) suffices, and the quotient is therefore non-local. This
accounts for the essential use of WISC in the construction of ordinals.

More generally, any quotient whose generators quantify over arbitrarily
deep substructures or require unbounded transitive closure will fall
outside the \(\alpha\)-local fragment.

\section{Method Overview}\label{method-overview}

Technically, we construct stratified diagrams indexed by ordinals below
\(\alpha\), and show that all cones factor through a bounded stage.
Plumpness of the ordinal ensures the existence of suprema for such
cones, yielding cocontinuity without choice.

\section{Related Work}\label{related-work}

Our work builds on the QW framework of Fiore, Pitts, and Steenkamp
\cite{fiore2022-quotients-inductive-types}, and complements the no-go
results of Shulman and Lumsdaine
\cite{lumsdaine2020-semantics-higher-inductive}. It is also related to
Pitts and Steenkamp's notion of infinitary inflation
\cite{steenkamp2021-phd-thesis}, which similarly isolates sources of
non-constructivity in inductive definitions.

\section{Future Work}\label{future-work}

Future directions include formalising the locality theorem stated above,
and extending the analysis to indexed QW types
\cite{altenkirch2018-quotient-inductive-inductive}.

\section{Conclusion}\label{conclusion}

We identify bounded locality as a structural criterion characterising
which quotient-W types admit fully constructive constructions. This
explains when choice principles such as WISC are genuinely necessary,
and when they can be avoided. Our results show that many infinitary
quotients previously thought to require choice are, in fact,
constructible in a purely constructive setting.

\section{Major Revision 2: After rewrite (7bc678e) - Streamlined version focusing on mobiles}

\section{Motivation}\label{motivation}

Quotient Inductive Types (QITs) are a central mechanism for specifying
datatypes equipped with identifications in set-truncated Homotopy Type
Theory \cite{hott2013-book}. They support a wide range of fundamental
constructions, including the HoTT reals \cite{hott2013-book}, partiality
monads \cite{altenkirch2017-partiality-monad}, and ordinal-like
structures \footnote{TODO: Citaiton}.

Despite their expressive power, QITs present significant foundational
difficulties. Shulman and Lumsdaine established a no-go theorem showing
that certain QITs, including Brower's standard definition of the countable
ordinals, cannot be constructed using quotients alone
\cite{lumsdaine2020-semantics-higher-inductive}. This suggests that some
form of infinitary principle or choice is unavoidable in a constructive
metatheory, at least in the general case.

Fiore, Pitts, and Steenkamp subsequently showed that a broad class of
QITs, which they call Quotient-W Types (QW types), admit an Initial
Algebra semantics assuming the Weak Initial Set of Covers principle (WISC)
\cite{fiore2022-quotients-inductive-types}. Their result
recovers many important examples, including the extensional countable
ordinals. However, WISC remains a non-constructive choice principle
\cite{vandenberg2012-wisc-axiom}, and its necessity is poorly
understood. In particular, it is unclear which features of a QIT
genuinely require choice, and which arise from limitations of existing
proof techniques.

This leads to our central research question:

\begin{quote}
	Which quotient-W types can be constructed fully constructively, without
	assuming any form of choice?
\end{quote}

%TODO: Define point and path constructor in general?
%TODO: Reword?
Our approach is to study a 'just-tricky-enough' example to test the need
for WISC in defining QITs, while still being as simple as possible. We then try
to prove that these exhibit an $(Q,sup)$ initial algebra---a categorical structure that
represents that there is exists a recursor (morphism between algebras), and that
that recursor is unique (no `extra stuff' in Q beyond what is needed to
represent the signature). Our plan was to exploritoriliy formalize it in Agda,
until we got stuck, and then to see exactly why WISC was needed in this case,
and see if we could determine a weaker principle, that was more justified
constructively. What we discovered, to our surprise, is that the step that Fiore et al. used WISC
could actually be shown to be fully constructive. We postulate that a large
class of QITs can be exhibited constructively.

%TODO: Define WISC?

\section{Mobiles}
The example that we chose--`mobiles'--have just two point constructors and one
path constructors, making it much simpler than the other 'hard' cases we
surveyed, while still being considered as challenging from a constructability
standpoint.

We define *mobiles* as well-founded trees quotiented by permutation at each
level. (TODO: Figures). Specifically,

$$
\begin{array}{l@{\; }l@{\quad}l}
  \multicolumn{3}{l}{\textbf{data } \mathsf{Mobile} : \mathcal{U} \textbf{ where}} \\
  \quad \mathsf{leaf}        & : \mathsf{Mobile} \\
  \quad \mathsf{node}(f)     & : \mathsf{Mobile}                                  & \forall f : \mathsf{Mobile}^I \\
  \quad \mathsf{perm}(\pi,f) & : \mathsf{node}(f) = \mathsf{node}(f \circ \pi) & \forall \pi : I \cong I ;\, f : \mathsf{Mobile}^I
\end{array}
$$

for some $I:\mathcal{U}$ where $\mathsf{Mobile}^I$ represents functions from the
branching index set $I$ and $\mathsf{Mobile}$'s. This has a clean, natural
presentation that fits in three lines, unlike others which require auxillary
data types, or stronger recursion semantics. %TODO: Terminology?

\section{Main Idea}\label{main-idea}

We briefly recall the framework of quotient-W types introduced by Fiore,
Pitts, and Steenkamp \cite{fiore2022-quotients-inductive-types}. A QW
signature consists of a container (representing branching structure) and a set
of equations over the container. From such a signature, Fiore et al.~construct a diagram of
approximations indexed by a size structure
\cite{fiore2022-quotients-inductive-types}. Intuitively, this diagram
represents successive stages of the quotient construction, where each
stage gives the proofs available within a bounded depth. The desired QW
type is obtained as the colimit $Q$ of this diagram.

Fiore et al.~show this that this diagram is in fact an initial algebra: having a
unique recursor, thereby showing that the structure using (indexed) WISC
\cite{fiore2022-quotients-inductive-types}. In particular,
WISC is crucial in proving *cocontinuity*, this case meaning that the
constructors lift to the quotiented form, and that $Q$ is closed under all
constructors.

The proof in Fiore et al.~relies on WISC to select a cover to size-bound the
family of subtrees, enabling the construction of inverse maps witnessing
this isomorphism.

Our work revisits this step. We observe that the need for WISC arises
precisely when the equations allow arbitrarily deep inspection of terms,
as in the extensional ordinal example. When all equations are
depth-bounded, the required coherence data can instead be constructed by
well-founded recursion on a fixed ordinal bound \(\alpha\), avoiding any
appeal to choice.

%TODO: Cut?
%We formalise the notion of locality by assigning a \emph{rank} function
%
%\[ \mathrm{rank} : T_\Sigma X \to \mathcal{O} \]
%
%measuring the height of the underlying tree. An equation system is said
%to be \(\alpha\)-local if, for every equation \(e : E\), both sides
%\(l(e)\) and \(r(e)\) have rank \emph{at most} a fixed ordinal
%\(\alpha\). Intuitively, this means that every generating equation only
%inspects structure up to depth \(\le \alpha\).
%
%Our main theorem is that \(\alpha\)-local QW signatures admit fully
%constructive initial algebra semantics.

%\emph{Theorem} Let \(\Sigma\) be a QW signature whose equations are
%\(\alpha\)-local. Then the associated quotient-W type exists as an
%initial algebra, and its construction does not require WISC or any other
%choice principle.
%
%The proof follows the outline of Fiore--Pitts--Steenkamp
%\cite{fiore2022-quotients-inductive-types}, but replaces the use of WISC
%in the inverse map by offsetting by a bound ordinal \(\alpha\), giving a
%constructable function between term bound and maximum proof depth,
%giving cocontinuity. This construction is fully constructive and does
%not require any choice principles.
%
%Our analysis explains why certain examples genuinely require choice. In
%the case of extensional ordinals, equality is defined by mutual
%containment of downward closures, which requires unbounded descent into
%subtrees \cite{fiore2022-quotients-inductive-types}. No ordinal bound
%\(\alpha\) suffices, and the quotient is therefore non-local. This
%accounts for the essential use of WISC in the construction of ordinals.
%
%More generally, any quotient whose generators quantify over arbitrarily
%deep substructures or require unbounded transitive closure will fall
%outside the \(\alpha\)-local fragment.
%
%Technically, we construct stratified diagrams indexed by ordinals below
%\(\alpha\), and show that all cones factor through a bounded stage.
%Plumpness of the ordinal ensures the existence of suprema for such
%cones, yielding cocontinuity without choice.

\section{Conclusion}\label{conclusion}

We identify bounded locality as a structural criterion characterising
which quotient-W types admit fully constructive constructions. This
explains when choice principles such as WISC are genuinely necessary,
and when they can be avoided. Our results show that many infinitary
quotients previously thought to require choice are, in fact,
constructible in a purely constructive setting.

Our work builds on the QW framework of Fiore, Pitts, and Steenkamp
\cite{fiore2022-quotients-inductive-types}, and complements the no-go
results of Shulman and Lumsdaine
\cite{lumsdaine2020-semantics-higher-inductive}. It is also related to
Pitts and Steenkamp's notion of infinitary inflation
\cite{steenkamp2021-phd-thesis}, which similarly isolates sources of
non-constructivity in inductive definitions.

Future directions include formalising the locality theorem stated above,
and extending the analysis to indexed QW types
\cite{altenkirch2018-quotient-inductive-inductive}.




\begin{thebibliography}{10}

\bibitem{altenkirch2016-qit}
Thorsten Altenkirch and Ambrus Kaposi.
\newblock Type theory in type theory using quotient inductive types.
\newblock In {\em Proceedings of the 43rd ACM SIGPLAN-SIGACT Symposium on
  Principles of Programming Languages (POPL 2016)}, pages 18--29, 2016.

\bibitem{altenkirch2018-qiit}
Thorsten Altenkirch, Paolo Capriotti, Gabe Dijkstra, Nicolai Kraus, and
Fredrik Nordvall Forsberg.
\newblock Quotient inductive-inductive types.
\newblock In {\em Foundations of Software Science and Computation Structures
  (FoSSaCS 2018)}, pages 293--310, 2018.

\bibitem{fiore2022-qwit}
Marcelo P.~Fiore, Andrew M.~Pitts, and S.~C.~Steenkamp.
\newblock Quotients, inductive types, and quotient inductive types.
\newblock {\em Logical Methods in Computer Science}, 18(2), 2022.

\bibitem{kaposi2019-constructing}
Ambrus Kaposi, Andr\'as Kov\'acs, and Thorsten Altenkirch.
\newblock Constructing quotient inductive-inductive types.
\newblock {\em Proceedings of the ACM on Programming Languages},
  3(POPL):2:1--2:24, 2019.

\bibitem{kovacs2020-large}
Andr\'as Kov\'acs and Ambrus Kaposi.
\newblock Large and infinitary quotient inductive-inductive types.
\newblock In {\em Proceedings of the 35th Annual ACM/IEEE Symposium on Logic in
  Computer Science (LICS 2020)}, pages 648--661, 2020.

\bibitem{kaposi2021-setoid}
Ambrus Kaposi and Zongpu Xie.
\newblock Quotient inductive-inductive types in the setoid model.
\newblock In {\em 27th International Conference on Types for Proofs and
  Programs (TYPES 2021)}, presentation abstract/slides, 2021.

\bibitem{lumsdaine2020-hit}
Peter LeFanu Lumsdaine and Michael Shulman.
\newblock Semantics of higher inductive types.
\newblock {\em Mathematical Proceedings of the Cambridge Philosophical
  Society}, 169(1):159--208, 2020.

\end{thebibliography}

\end{document}
